Distributed Multi-Heuristic A* (HAMSTAR)

### 2014-12-11 paper.tex
Results solidification

Figure \ref{fig1} shows performance as we increase the number of processes (or heuristics, in the case of SMHA*). As shown, performance initially increases significantly, but then drops off as the number of processes increases. When running on a single machine, our distributed algorithm performs comparably to SMHA* for 4 or fewer cores, but decreases in performance as more processes are added.

There are a number of likely explanations for this. First, the initial performance increase is largely a result of adding a heuristic which has more flexibility in returning a less-optimal path. Indeed, figure \ref{fig2} shows that the solution length increases considerably after adding just one non-anchor process. However, it is worth noting that our goal is specifically defined as trying to find a path within a specific multiple of the optimum. Thus, a longer path that remains within these bounds is viewed as an equally valid solution. The additional heuristics also provides more variety in the search, so that it is more likely that at least one of them succeeds in finding a good path. The drop in performance as more processes are added appears to be due to the increased communication cost of using a broadcast communication system. All processes must pause for this, and some might be slow in calling the broadcast command. Moreover, as more processes are added, the anchor receives more states that it must sift through, and the following message that it broadcasts increases in size as well. Thus, it appears that after a sufficiently diverse group of heuristics is added, the added communication cost of additional heuristics outweighs the benefits. These communication costs are analyzed in more detail later in this section and displayed in figure \ref{fig5}.

Figure \ref{fig3} shows performance as we vary the number of iterations between communications. For reference, 100 iterations in our experiments is roughly 1-5 milliseconds. As one would expect, very frequent communication eventually causes the communication overhead to grow too large and hurts performance.  On the other hand, as communication becomes extremely infrequent, the lack of shared information also hurts performance considerably. The optimal choice therefore depends heavily on the nature of the problem; in our case, it appears to be around 6,400 iterations.

As discussed earlier, our algorithm does not appear to outperform SMHA* in the sliding tile puzzle domain. The most likely explanation for this is that because the heuristics we used for the puzzle can be calculated extremely efficiently, there are a huge number of expansions per unit of time. Each expanded state must be communicated to the anchor, which involves a considerable amount of work. It is likely that the amount of work in sending the states to the anchor, having the anchor process these states, broadcasting the best of those states to the other children, and having those children process the states, all together takes about as much work as expanding the state alone. Indeed, figure \ref{fig5} shows that as the number of processes increases, communication quickly becomes a huge fraction of the total time spent. This increased communication cost ultimately outweighs the benefits of additional processes.

Nevertheless, figure \ref{fig4} displays what the results might look like in a problem where the heuristic takes longer to calculate. Specifically, we artificially slow the heuristic calculation by a factor of 10. The result is that fewer states are expanded per second, and thus the cost of communicating a state is much lower relative to the cost of expanding it. In this case, the results show that our distributed algorithm can leverage the benefit of additional cores to outperform SMHA* if we use the one-on-one communication method described earlier in this paper.

@@

>>>>

Figure \ref{fig1} shows performance as we increase the number of processes (or heuristics, in the case of SMHA*). As shown, performance increases significantly at first, but then drops off as the number of processes increases. When running on a single machine, our distributed algorithm performs comparably to SMHA* for 4 or fewer cores, but decreases in performance as more processes are added.

There are a number of likely explanations for this. First, the initial performance increase is largely a result of adding a heuristic with a greedier bias, typically finding a slightly less optimal path more quickly. Indeed, figure \ref{fig2} shows that the solution length increases considerably after adding just one non-anchor process. However, it is worth noting that our criterion is specifically to find a path whose cost is within the specified multiple (in our case, 4) of the optimum. Crucially, the combination of monotone and inadmissible search is able to guarantee the near-optimality of the solution; thus, the slightly longer path is viewed as an equally valid solution. The additional heuristics provide more variety in the search, making it more likely for at least one of them (or some combination) to succeed in finding a good path. The drop in performance as more processes are added appears to be due to the increased communication cost of using a broadcast communication system. All processes must pause for this, and some might be slow in calling the broadcast command. Moreover, as more processes are added, the anchor receives more states that it must sift through, and the following message that it broadcasts increases in size as well. Thus, it appears that after a sufficiently diverse group of heuristics is added, the added communication cost of additional heuristics outweighs the benefits. These communication costs are analyzed in more detail later in this section and are displayed in figure \ref{fig5}.

Figure \ref{fig3} shows performance as we vary the number of iterations between communications. For reference, 100 iterations in our experiments takes roughly 1-5 milliseconds. As one would expect, very frequent communication eventually causes the communication overhead to grow too large and hurts performance.  On the other hand, as communication becomes extremely infrequent, the lack of shared information also hurts performance considerably. The optimal communication interval therefore depends heavily on the nature of the problem; in our case, it appears to be around 6,400 iterations.

As discussed earlier, our algorithm does not appear to outperform SMHA* in the sliding tile puzzle domain. The most likely explanation for this is that, because the successor states and heuristics we used for the puzzle can be calculated extremely efficiently, there are a huge number of expansions per unit time. Each expanded state must be communicated to the anchor, which involves a considerable amount of work. It is likely that the amount of work in sending the states to the anchor, having the anchor process these states, broadcasting the best of those states to the other children, and having those children process the states, altogether takes about as much work as expanding the state alone. Indeed, figure \ref{fig5} shows that as the number of processes increases, communication quickly becomes a huge fraction of the total time spent. This increased communication cost ultimately outweighs the benefits of additional processes.

Nevertheless, figure \ref{fig4} displays what the results might look like in a problem where the heuristic takes longer to calculate. Specifically, we artificially slow the heuristic calculation by a factor of 10. The result is that fewer states are expanded per second, and thus the cost of communicating a state is much lower relative to the cost of expanding it. In this case, our distributed algorithm can leverage the benefit of additional cores to outperform SMHA* if we use the one-on-one communication method described earlier in this paper. Thus, a distributed approach may prove helpful in domains such as manipulator arm planning, in which expansions are slowed by the collision checking and numerical integration needed to compute successor states and heuristics.

@@

\begin{figure}
\centering
\includegraphics[width=1.6in]{graphs/hamstar}
\end{figure}
